---
title: "anova model checking"
author: "Saima Jahan"
date: "10/21/2020"
output:
  word_document: default
  html_document:
    df_print: paged
---
Step 1: Load the data into R


```{r}
getwd()
setwd("/spring 2020/ast -504/lec")
getwd()
dat<-read.csv("plasma.csv")
dat

```

Step 2: Check that the data meets the assumptions

a. The errors are assumed to be normally distributed with mean zero.



```{r}
library(ggplot2)
h1<-ggplot(dat, aes(x=etch.rate)) + 
  geom_histogram( breaks=seq(530,725,by=30), 
                 col="black", 
                 fill="blue",alpha=0.4
                 ) + 
  labs(title="Histogram ", x="Class", y="Frequency") 
# Change the line type and color of axis lines
h1<-h1+ theme( axis.line = element_line(colour = "darkblue", 
                      size = 1, linetype = "solid"))+theme(plot.title = element_text(hjust = 0.5))

 
h1
```

Create Q-Q plot of residuals

```{r}
#create Q-Q plot to compare this dataset to a theoretical normal distribution 
qqnorm(dat$etch.rate)

#add straight diagonal line to plot
qqline(dat$etch.rate)
```



# Conduct Shapiro-Wilk Test for Normality
 
null: samples come from a normal distribution
 
```{r}
#Conduct Shapiro-Wilk Test for normality 
shapiro.test(dat$etch.rate)

```

b. Equal Variance


```{r}
library(ggplot2)
box<-ggplot(dat,aes(x=etch.rate,y=treatment))+geom_boxplot(fill = "blue",col="black",alpha=0.4)+theme( axis.line = element_line(colour = "darkblue", 
                      size = 1, linetype = "solid"))+labs(title="Check the assumption visually using boxplots ", x="Etch rate",y="Treatment")+theme(plot.title = element_text(hjust = 0.5))
box

```

Conduct Bartlett’s Test

null hypothesis: The samples have equal variances.


```{r}
bartlett.test(etch.rate ~ treatment, data=dat)
```

c. Independence

The observations in each group are independent of the observations in every other group.

```{r}
anova<- aov(etch.rate ~ treatment ,data=dat)
anova
summary(anova)


plot(anova)
```
Fitted versus residual plot:
This shows if there is a pattern in the residuals, and ideally should show similar scatter for each condition.
Fitted versus root over standardized residual:
This is like the first plot but now to specifically test if the residuals increase with the fitted values.


lec 2 fall 20

```{r}
treatment<-c("20g","20g","20g","20g","30g","30g","30g","30g","40g","40g","40g","40g")
length(treatment)




observation<-c(24,28,37,30,37,44,39,35,42,47,52,38)
length(observation)


df<-data.frame(treatment,observation)
df
ano<-aov(df)
```


```{r}
getwd()
setwd("/Fall 20/AST 504/question/")
getwd()

```


```{r}
getwd()
setwd()
```
shozib s analysis

16.11.20 

```{r}

```


lec 5 

##read data

data<-read.csv(file.choose(), header=TRUE)
data

  obs Method
1   3.34      1
2   3.12      1
3   4.23      1
4   3.70      1
5   4.75      1
6   3.12      1
7   3.91      2
8   5.94      2
9   5.14      2
10  5.36      2
11  5.86      2
12  7.55      2
13  9.31      3
14 11.37      3
15 12.75      3
16  9.09      3
17 12.82      3
18 10.24      3
19 20.15      4
20 14.82      4
21 13.95      4
22 20.20      4
23 17.35      4
24 19.82      4


## Model adequacy plot

MA<-as.factor(data$Method)
MA

crd<-aov(data$obs~MA, data=data)
crd

par(mfrow=c(2,2))
plot(crd)


##variance Test for checking equality of variance on the raw data


bartlett.test(data$obs~data$Method)


        Bartlett test of homogeneity of variances

data:  data$obs by data$Method
Bartlett's K-squared = 8.9958, df = 3, p-value = 0.02935

##transformation method 

meann<-tapply(data$obs,data$Method, mean)
meann

##Mean

        1         2         3         4 
 3.710000  5.626667 10.930000 17.715000 

##standard deviation

sdd<-tapply(data$obs,data$Method, sd)
sdd

       1        2        3        4 
0.661090 1.192202 1.647070 2.800891 

##log of mean

lmeu<-read.table("clipboard")
lmeu

         V1
1 0.5693739
2 0.7502512
3 1.0386202
4 1.2483412

##log of standard deviation

ls<-read.table("clipboard")
ls

         V1
1 -0.17973941
2  0.07634985
3  0.21671206
4  0.44729621

##regression Analysis

reg<-lm(ls$V1~lmeu$V1)
reg


Call:
lm(formula = ls$V1 ~ lmeu$V1)

Coefficients:
(Intercept)      lmeu$V1  
    -0.6311       0.8554 


## scatter plot for cheaking straight line with slope
 
plot(ls$V1,lmeu$V1)


## apply lognormal transformation method

data1<-read.table("clipboard")
data1

          V1 V2
1  0.5237465  1
2  0.4941546  1
3  0.6263404  1
4  0.5682017  1
5  0.6766936  1
6  0.4941546  1
7  0.5921768  2
8  0.7737864  2
9  0.7109631  2
10 0.7291648  2
11 0.7678976  2
12 0.8779470  2
13 0.9689497  3
14 1.0557605  3
15 1.1055102  3
16 0.9585639  3
17 1.1078880  3
18 1.0103000  3
19 1.3042750  4
20 1.1708482  4
21 1.1445742  4
22 1.3053514  4
23 1.2392995  4
24 1.2971036  4


## Model adequacy Plots on the transformed data:

F<-as.factor(data1$V2)
F

crd1<-aov(data1$V1~F)
crd1
par(mfrow=c(2,2))
plot(crd1)


## Again variance test on the transformed data for cheking variance equality


bartlett.test(data1$V1~data1$V2)


        Bartlett test of homogeneity of variances

data:  data1$V1 by data1$V2
Bartlett's K-squared = 0.66993, df = 3, p-value = 0.8803



##anova test


MA<-as.factor(data1$V2)
MA

anova<-aov(data1$V1~MA)
anova

Call:
   aov(formula = data1$V1 ~ MA)

Terms:
                       MA Residuals
Sum of Squares  1.6440679 0.1187223
Deg. of Freedom         3        20

Residual standard error: 0.07704619
Estimated effects may be unbalanced

summary(anova)


            Df Sum Sq Mean Sq F value   Pr(>F)    
MA           3 1.6441  0.5480   92.32 6.88e-12 ***
Residuals   20 0.1187  0.0059                     
---
Signif. codes:  0 â***â 0.001 â**â 0.01 â*â 0.05 â.â 0.1 â â 1

## Comparisons among treatment means:

TukeyHSD(anova)

  Tukey multiple comparisons of means
    95% family-wise confidence level

Fit: aov(formula = data1$V1 ~ MA)

$MA
         diff        lwr       upr     p adj
2-1 0.1781074 0.05360332 0.3026115 0.0035830
3-1 0.4706135 0.34610940 0.5951175 0.0000000
4-1 0.6796934 0.55518937 0.8041975 0.0000000
3-2 0.2925061 0.16800202 0.4170102 0.0000117
4-2 0.5015860 0.37708198 0.6260901 0.0000000
4-3 0.2090800 0.08457589 0.3335840 0.0007323
```{r}

getwd()
s<-read.csv("assgmnt 504 asma.csv")
```

```{r}

str(s)

```
```{r}
crd<- aov(dep ~ 
         
               Method,data=s)
summary(crd)
plot(crd)
par(mfrow=c(2,2))
plot(crd)

```
```{r}
bartlett.test(dep~Method,data=s)

```
```{r}
tapply(s$dep,s$Method, mean) 


tapply(s$dep, s$Method,sd) 
```
```{r}
sigmayi<-c(.66,1.19,1.65,2.80)
logsigmayi<-log(sigmayi)
logsigmayi
```
```{r}
mui<-c(.71,2.63,7.93,14.72)
logmui<-log(mui)
logmui
```
```{r}
lm(logsigmayi~logmui)
```
```{r}
s<-read.csv("assgmnt 504 asma.csv")
```

```{r}
crd<- aov(dep1 ~ 
         
               Method,data=s)
summary(crd)
plot(crd)
par(mfrow=c(2,2))
plot(crd)


```
```{r}
bartlett.test(dep1~Method,data=s)
```
```{r}

crd<- aov(s$dep1 ~ 
             Method,data=s)
 
 par(mfrow=c(2,2))
 plot(crd)
 a<-bartlett.test(s$dep1~s$Method)
 a
 
 summary(crd)
```

```{r}
plot(logmui,logsigmayi,type = "l")
```
2.12.20
```{r}
etchrate<-c(575,542,530,539,570,565,593,590,579,610,600,651,610,637,629,725,700,715,685,710)
treatment<-c("RF160","RF180","RF200","RF220")
df<-data.frame(etchrate,treatment)
df

View(df)
```


another 

```{r}
etchrate1<-c(575,542,530,539,570,565,593,590,579,610,600,651,610,637,629,725,700,715,685,710)
treatment1<-c("RF160","RF160","RF160","RF160","RF160","RF180","RF180","RF180","RF180","RF180","RF200","RF200","RF200","RF200","RF200","RF220","RF220","RF220","RF220","RF220")
df1<-data.frame(etchrate1,treatment1)
df1
View(df1)
```






```{r}
etchrate<-c(575,565,600,725,542,593,651,700,530,590,610,715,539,579,637,685,570,610,629,710)
treatment<-c("RF160","RF180","RF200","RF220")
df<-data.frame(etchrate,treatment)
df
model<-aov(df$etchrate~df$treatment)
summary(model)
View(df)
```

```{r}
sessionInfo()
```
```{r}
?aov
```


```{r}
?anova
```
